---
title: "lrem-Introduction"
author: "Tatsuya Asami"
date: '`r format(Sys.time(), "%Y/%m/%d")`'
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{lrem-Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This source simulates a linear ratinal expectation model as following,

$$
E \mathbb{E}_t x_{t+1} = A x_{t} + B u_{t}, \ \ x_{t} \in \mathbb{R}^{n}, \ \ u_{t} \in \mathbb{R}^{m}.
$$
$E$ need not be a reversible matrix. $x_{t} = (x^{1}_{t}, \ x^{2}_{t})^{t}$ and $x^{1}_{t}$ is the predetermined vector, $x^{2}_{t}$ is the un-predetermined vector. There must exist some z such that $\det(E z - A) \neq 0$. 

We transform this system into the homogeneous system when the shock follows the AR process $u_{t + 1} = \Phi u_{t} + \epsilon_{t + 1}$

$$
\begin{bmatrix} I & O \\ O & E \end{bmatrix} \mathbb{E}_t \begin{bmatrix} u_{t + 1} \\ x_{t + 1} \end{bmatrix} = 
\begin{bmatrix} \Phi & O \\ B & A \end{bmatrix} \begin{bmatrix} u_{t} \\ x_{t} \end{bmatrix}
$$

## Example 1 (Hansen's RBS model)
We replicate the example in https://cdn.rawgit.com/rokko-ed17q1/final/64730073/lrem-hansen-rbc.html.

The parameters are

```{r}
alpha = 0.33
beta = 0.99
delta = 0.023
chi = 1.75
rho = 0.95
```

```{r}
q0 = (1 - beta + beta * delta) / alpha / beta
q1 = q0 ^ (1 / (1 - alpha))
q2 = q0 - delta

kbar = (1 - alpha) * q1 ^ (- alpha)
kbar = kbar / ((1 - alpha) * q0 + chi * q2)

cbar = q2 * kbar
nbar = q1 * kbar
zbar = 1
```

The matrices are

```{r}
E = matrix(0, 3, 3)
A = matrix(0, 3, 3)
B = matrix(0, 3, 1)
Phi = matrix(rho, 1, 1)

E[1, 1] = alpha * (alpha - 1) * q0
E[1, 2] = alpha * q0
E[1, 3] = - (1 - delta + alpha * q0)
E[2, 1] = 1

A[1, 3] = E[1, 3]
A[2, 1] = - A[1, 3]
A[2, 2] = (1 - alpha) * q0
A[2, 3] = - q2
A[3, 1] = alpha
A[3, 2] = (- alpha - (1 - alpha) * nbar) / (1 - nbar)
A[3, 3] = -1

B[1, 1] = - alpha * q0 * rho
B[2, 1] = q0
B[3, 1] = 1
```

The simulation is following

```{r}
library("lrem")
library("ggplot2")
```

```{r}
policy <- lre_ar(A, E, B, Phi, 0, 1)

# Simulation period
steps <- 100

# Path of the growth rates
out <- simulate(policy$g, policy$h, 0, 1, steps) 
out0 <- matrix(0, nrow = 1, ncol = ncol(out))
out <- rbind(out0, out)
# Steady State
ss <- c(zbar, kbar, nbar, cbar)

# Convert percentage changes to levels
for (i in 1:ncol(out)) {
  out[, i] <- ss[i] * out[, i] + ss[i]
}
```

```{r}
result <- data.frame(out)
names(result) <- c("Z", "K", "N", "C")
result["t"] <- 0:(steps)
```

```{r}
# Plotting
ggplot(result) + geom_line(aes(x = t, y = Z))
```

```{r}
ggplot(result) + geom_line(aes(x = t, y = K))
```

```{r}
ggplot(result) + geom_line(aes(x = t, y = N))
```

```{r}
ggplot(result) + geom_line(aes(x = t, y = C))
```

## Example 2 (the canonical New Keynesian model)

We simulate the following New Keynesian model.

*Phillips curve*

$$
\pi_{t} = \beta \mathbb{E}_{t} \pi_{t + 1} + \kappa x_{t} + u^{S}_{t}
$$

*IS curve*

$$
x_{t} = \mathbb{E}_{t} x_{t + 1} - \frac{1}{\sigma} (i_{t} - \mathbb{E}_{t} \pi_{t + 1}) + u^{D}_{t}
$$

*Monetary policy rule*

$$
i_{t} = \alpha \pi_{t} + \iota x_{t} 
$$

Assume AR shock with $|\rho_{S}|, |\rho_{D}|  < 1$ and $\mathbb{E}_{t} \epsilon^{S}_{t + 1} = \mathbb{E}_{t} \epsilon^{D}_{t + 1}$;

$$
u^{S}_{t + 1} = \rho_{S} u^{S}_{t} + \epsilon^{S}_{t + 1} \\
u^{D}_{t + 1} = \rho_{D} u^{D}_{t} + \epsilon^{D}_{t + 1}
$$

  - $\pi_{t}$: log inflation rate
  - $x_{t}$: output gap
  - $i_{t}$: log gross nominal interest rate
  - $\sigma$: intertemporal elastocity of substitution
  - $\beta$: discount rate
  - $\kappa$: slope pf the Phillips curve
  - $\alpha$ and $\iota$: monetary policy parameters
  
Rewriting this system into the above matrix form, we obtain

$$
\begin{bmatrix} I & O \\ O & E \end{bmatrix} \mathbb{E}_t \begin{bmatrix}  u_{t + 1} \\ \pi_{t + 1} \\ x_{t + 1} \\ i_{t + 1} \end{bmatrix} = 
\begin{bmatrix} \Phi & O \\ B & A \end{bmatrix} \begin{bmatrix} u_{t} \\ \pi_{t} \\ x_{t} \\ i_{t} \end{bmatrix}
$$
where

$$
A = \begin{bmatrix} 1 & - \kappa & 0 \\ 0 & 1 & 1 / \sigma \\ \alpha & \iota & - 1 \end{bmatrix}, \ 
E = \begin{bmatrix} \beta & 0 & 0 \\ 1 / \sigma & 1 & 0 \\ 0 & 0 & 0 \end{bmatrix}, \ 
B = \begin{bmatrix} - 1 & 0 \\ 0 & - 1 \\ 0 & 0 \end{bmatrix}, \\
\Phi = \begin{bmatrix} \rho_{S} & 0 \\ 0 & \rho_{D} \end{bmatrix}, \ 
u_{t} = \begin{bmatrix} u^{S}_{t} \\ u^{D}_{t} \end{bmatrix}.
$$


The parameters are

```{r}
alpha <- 2
iota <- 0
beta <- 0.99
sigma <- 1
kappa <- 0.132
rhoS <- 0.9
rhoD <- 0.9
```

The matrices are
```{r}
A <- matrix(c(1, - kappa, 0,
              0, 1, 1 / sigma,
              alpha, iota, - 1), nrow = 3, byrow = TRUE)
E <- matrix(c(beta, 0, 0,
              1 / sigma, 1, 0,
              0, 0, 0), nrow = 3, byrow = TRUE)
B <- matrix(c(- 1, 0,
              0, - 1,
              0, 0), nrow = 3, byrow = TRUE)
Phi <- matrix(c(rhoS, 0,
              0, rhoD), nrow = 2, byrow = TRUE)
```

The values of these variables at the steady state are

```{r}
rhoSbar <- 0
rhoDbar <- 0
pibar <- 0
xbar <- 0
ibar <- 0
ss <- c(rhoSbar, rhoDbar, pibar, xbar, ibar)
```

That's all of preparation of simulation.

```{r}
# Initial shock is (us, ud) = (1, 1)
policy <- lre_ar(A, E, B, Phi, NULL, c(1, 1))
# Simulation period
steps <- 100

out <- simulate(policy$g, policy$h, NULL, c(1, 1), steps)
out0 <- matrix(ss, nrow = 1, ncol = ncol(out))
out <- rbind(out0, out)
```

```{r}
result <- data.frame(out)
names(result) <- c("rhoS", "rhoD", "pi", "x", "i")
result["t"] <- 0:(steps)
```

```{r}
# Plotting
ggplot(result) + geom_line(aes(x = t, y = rhoS))
```


```{r}
ggplot(result) + geom_line(aes(x = t, y = rhoD))
```

```{r}
ggplot(result) + geom_line(aes(x = t, y = pi))
```

```{r}
ggplot(result) + geom_line(aes(x = t, y = x))
```

```{r}
ggplot(result) + geom_line(aes(x = t, y = i))
```






